---
title: "PQHS 602 - TableOne"
author: "Manu Bulusu"
date: last-modified
format: 
  html:
    toc: true
    number-sections: true
    code-fold: show
    code-tools: true
    code-overflow: wrap
    embed-resources: true
    date-format: iso
    theme: lux
---

## R Packages and Setup {.unnumbered}

```{r}
#| message: false
#| warning: false

knitr::opts_chunk$set(comment = NA) 
options(error = recover)

library(janitor) 
library(naniar)

library(broom)
library(caret)
library(car)
library(codebookr)
library(mice)
library(gt)
library(patchwork)
library(rms)
library(mosaic)
library(haven)
library(readr)
library(transplantr)
library(tableone)
library(table1)
library(nhanesA)
library(DescTools)
library(webshot2)
library(htmlwidgets)

library(easystats)
library(tidyverse) 

library(conflicted)

theme_set(theme_bw()) 
```

```{r}

conflicted::conflicts_prefer(
  dplyr::select,
  dplyr::filter,
  dplyr::rename,
  Hmisc::`label<-`
)

```


In this file, I will create TableOnes from my cleaned dataset with NHANES subjects from 2003-2018.

# Load Data

```{r}
studydata <- readRDS("mb_602_data_final.Rds")
```

```{r}
glimpse(studydata)
```

Relabeling predictor variables for presentation in tables:

```{r}
studydata <- studydata |>
  mutate(
    mort_status = factor(mort_status, levels = c(1, 0), labels = c("Event", "No Event"))
  )

label(studydata$age) <- "Age (years)"
label(studydata$sex) <- "Sex"
label(studydata$ethnicity) <- "Ethnicity"
label(studydata$arsenic_total) <- "Total Urinary Arsenic (ug/L)"
label(studydata$arsenic_mma) <- "Urinary MMA (ug/L)"
label(studydata$arsenic_dma) <- "Urinary DMA (ug/L)"
label(studydata$arsenic_asb) <- "Urinary AsB (ug/L)"
label(studydata$acr) <- "Albumin-Creatinine Ratio (mg/g)"
label(studydata$eGFR) <- "eGFR (mL/min/1.73 m^2)"
label(studydata$highest_ed) <- "Highest Education"
label(studydata$smoking) <- "Smoker"
label(studydata$diabetes) <- "Diabetes"
label(studydata$hypertension) <- "Hypertension" 
label(studydata$bmi) <- "BMI (kg/m^2)"
label(studydata$urine_cr) <- "Urinary Creatinine (mg/dL)"
label(studydata$prostate_cancer) <- "Prostate Cancer"
label(studydata$mort_status) <- "All-Cause Mortality"
```

# Prostate Cancer

I will now make a Table1 stratified by each outcome, presenting all exposures and covariates.

First removing all female subjects from the sample:

```{r}
studydata_prostate <- studydata |>
  filter(sex == "Male") |>
  select(
    prostate_cancer, arsenic_total, arsenic_mma, arsenic_dma, arsenic_asb, acr, age, ethnicity, highest_ed, smoking, bmi, diabetes,
    hypertension, urine_cr, eGFR
    ) 
```

Now making the Table1 stratified by prostate cancer status:

```{r}
prostate_table1 <- table1(~ age + ethnicity + arsenic_total + arsenic_mma + arsenic_dma + arsenic_asb +
          highest_ed + smoking + diabetes + hypertension + bmi + acr + eGFR + urine_cr | prostate_cancer, data = studydata_prostate, caption = "Prostate Cancer")

```

```{r}
prostate_table1
```


# All-Cause Mortality

Now, I will look at all-cause mortality while retaining the same variables described in the cancer-outcome tables above:

```{r}
mort_table1 <- table1(~ age + sex + ethnicity + arsenic_total + arsenic_mma + arsenic_dma + arsenic_asb +
          highest_ed + smoking + diabetes + hypertension + bmi + acr + eGFR + urine_cr | mort_status, data = studydata, caption = "All-Cause Mortality")
```

```{r}
mort_table1
```

# Session Information

```{r}
xfun::session_info()
```
