---
title: "PQHS 602 - Codebook"
author: "Manu Bulusu"
date: last-modified
format: 
  html:
    toc: true
    number-sections: true
    code-fold: show
    code-tools: true
    code-overflow: wrap
    embed-resources: true
    date-format: iso
    theme: lux
---

## R Packages and Setup {.unnumbered}

```{r}
#| message: false
#| warning: false

knitr::opts_chunk$set(comment = NA) 
options(error = recover)

library(janitor) 
library(naniar)

library(broom)
library(caret)
library(car)
library(codebookr)
library(mice)
library(gt)
library(patchwork)
library(rms)
library(mosaic)
library(haven)
library(readr)
library(transplantr)
library(tableone)
library(nhanesA)
library(conflicted)
library(webshot2)

library(easystats)
library(tidyverse) 

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("rename", "dplyr")
conflict_prefer("max", "base")

theme_set(theme_bw()) 
```

# Load Data

```{r}
studysample <- readRDS("mb_602_data_final.Rds")
```

```{r}
glimpse(studysample)
```


# Codebook

| Variable | Role | Type | Description |
|-----------------:|:----------------:|:----------------:|:-----------------|
| `NHANES_ID` | identifier | \- | Unique subject identifier for unified NHANES data, consists of survey cycle appended to original NHANES id (SEQN) |
| `age` | predictor | quantitative | Subject age - *years* |
| `sex` | predictor | binary | Subject sex - *Male, Female* |
| `ethnicity` | predictor | multi-categorical | Subject ethnicity - *Mexican American, Other Hispanic, Non-Hispanic White, Non-Hispanic Black, Other Race - Including Multi-Racial* |
| `prostate_cancer` | outcome | binary | Has subject been diagnosed with prostate cancer - *Yes, No* |
| `mort_status` | outcome | binary | Subject all-cause mortality status, as reported in LMF data - *0 = Censored (Alive or Lost to Follow-up), 1 = Event Observed (Deceased)* |  
| `arsenic_total` | predictor | quantitative | Subject urinary arsenic, total - *ug/L* |
| `arsenic_mma` | predictor | quantitative | Subject urinary monomethylarsonic acid - *ug/L* |
| `arsenic_dma` | predictor | quantitative | Subject urinary dimethylarsonic acid - *ug/L* |
| `arsenic_asb` | predictor | quantitative | Subject urinary arsenobetaine - *ug/L* |
| `highest_ed` | predictor | multi-categorical | Subject highest education level achieved - *No High School Diploma, High School Graduate or GED, Some College or AA Degree, College Graduate* |
| `smoking` | predictor | binary | Has subject smoked at least 100 cigarettes in lifetime - *Yes, No* |
| `diabetes` | predictor | binary | Has subject been told they have diabetes by physician - *Yes, No* |
| `hypertension` | predictor | multi-categorical | Based on average blood pressure, does subject have hypertension - *Yes, No, Missing* |
| `bmi` | predictor | quantitative | Subject body mass index - *kg/m^2* |
| `acr` | predictor | quantitative | Subject albumin-creatinine ratio - *mg/g* |
| `eGFR` | predictor | quantitative | Subject estimated glomerular filtration rate - *mL/min/1.73 m^2* |
| `urine_cr` | predictor | quantitative | Subject urinary creatinine - *mg/dL* |
| `ELIGSTAT` | indicator | \- | Indicator of subject LMF eligibility, ensuring mortality data is available |
| `PERMTH_EXM` | outcome | quantitative | Time in person-months from subject examination to observed mortality event (specified by `mort_status`) |



# Numerical Description

```{r}
numerical_summary <- data_codebook(studysample) |>
  mutate(
    Description = case_when(
      Name == "NHANES_ID" ~ "Unique subject identifier for unified NHANES data, consists of survey cycle appended to original NHANES id (SEQN)",
      Name == "age" ~ "Subject age - years",
      Name == "sex" ~ "Subject sex",
      Name == "ethnicity" ~ "Subject ethnicity",
      Name == "prostate_cancer" ~ "Has subject been diagnosed with prostate cancer?",
      Name == "mort_status" ~ "Subject all-cause mortality status, as reported in LMF data - 0 = Censored (Alive or Lost to Follow-up), 1 = Event Observed (Deceased)",
      Name == "arsenic_total" ~ "Subject urinary arsenic, total - ug/L",
      Name == "arsenic_mma" ~ "Subject urinary monomethylarsonic acid - ug/L",
      Name == "arsenic_dma" ~ "Subject urinary dimethylarsonic acid - ug/L",
      Name == "arsenic_asb" ~ "Subject urinary arsenobetaine - ug/L",
      Name == "highest_ed" ~ "Subject highest education level achieved",
      Name == "smoking" ~ "Has subject smoked at least 100 cigarettes in lifetime?",
      Name == "diabetes" ~ "Has subject been told they have diabetes by physician?",
      Name == "hypertension" ~ "Based on average blood pressure, does subject have hypertension?",
      Name == "bmi" ~ "Subject body mass index - kg/m^2",
      Name == "acr" ~ "Subject albumin-creatinine ratio - mg/g",
      Name == "eGFR" ~ "Subject estimated glomerular filtration rate - mL/min/1.73 m^2",
      Name == "urine_cr" ~ "Subject urinary creatinine - mg/dL",
      Name == "ELIGSTAT" ~ "Indicator of subject LMF eligibility, ensuring mortality data is available",
      Name == "PERMTH_EXM" ~ "Time in person-months from subject examination to observed mortality event (specified by mort_status)",
      .default = ""
    )
  ) |>
  select(Name, Description, Type, Values, N, Prop, Missings) |>
  gt(rowname_col = "row") |>
  cols_align(
    align = "center",
    columns = c(Type, Missings, Values, N, Prop)
  ) |>
  opt_table_font(font = "Times New Roman", size = px(12))

numerical_summary
gtsave(numerical_summary, "numerical_summary.png")
```


# Session Information

```{r}
xfun::session_info()
```
